Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes

### 2020-02-19 ms/SI.rmd
Add SI table for spatial model of SLA

# Load analysis results
loadd(list = c(
  "taxonomic_data_for_si",
  "trait_data_for_si",
  "climate_ancova_results",
  "phylosig_binary_traits_strict",
  "correlated_evo_test_strict",
  "best_fit_div_models_moran",
  "func_comm_struc_combined",
  "climate_model_summaries",
  "climate_models",
  "div_el_models",
  "div_el_model_summaries",
  "cwsd_scatterplot"), 
  cache = canopy_cache)

@@

>>>>

# Load analysis results
loadd(list = c(
  "taxonomic_data_for_si",
  "trait_data_for_si",
  "climate_ancova_results",
  "phylosig_binary_traits_strict",
  "correlated_evo_test_strict",
  "best_fit_div_models_moran",
  "func_comm_struc_combined",
  "climate_model_summaries",
  "climate_models",
  "div_el_models",
  "div_el_model_summaries",
  "div_el_model_parameter_summaries",
  "cwsd_scatterplot"), 
  cache = canopy_cache)

@@

**`r s_table("spatial-sla")`** `r s_table_cap("spatial-sla")` (see separate file)

Spatial autocorrelation accounted for by including a Mátern spatial correlation structure in model (see Methods).

```{r spatial-sla}
# Format table
spatial_sla_formatted <-
div_el_model_parameter_summaries %>%
  filter(var == "sla") %>%
  select(term, Coefficient = estimate, SE = std.error, t = t_value) %>%
  mutate(term = term %>%
           str_replace_all("habitterrestrial", "Growth habit") %>%
           str_replace_all("el", "Elevation")) %>%
  rename(Term = term)

# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("spatial-sla") %>% wrap_quotes
)

footer <- "Spatial autocorrelation accounted for by including a Mátern spatial correlation structure in model (see Methods)." %>%
  str_wrap(width = 100) %>%
  str_split("\\n") %>%
  unlist %>%
  purrr::map_chr(wrap_quotes)

# Write out table with metadata
filename <- s_table("spatial-sla") %>%
  str_replace_all(" ", "_") %>%
  paste0(".csv")

# Write out table with metadata
write_csv_with_meta(
  spatial_sla_formatted, 
  here::here("results", filename), 
  header,
  footer)
```

### 2020-02-19 ms/manuscript.Rmd
Add SI table for spatial model of SLA

# Load analysis results
loadd(list = c(
  "best_fit_div_models",
  "climate_ancova_results",
  "climate_plot",
  "climate_raw",
  "climate_select",
  "combined_comm_div_plots",
  "combined_cwm_plots",
  "comm",
  "correlated_evo_test",
  "correlated_evo_test_gemmae",
  "daily_mean_climate",
  "div_el_model_summaries",
  "div_t_test_results",
  "div_metrics_all",
  "fern_traits",
  "importance_heatmap",
  "moorea_climate",
  "pca_plot",
  "pca_results",
  "phy",
  "phylosig_binary_traits",
  "phylosig_cont_traits",
  "pic_results",
  "species_list",
  "traits_with_tree"),
  cache = canopy_cache)

@@

We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, 'climatic variables').
Variables relating to VPD, which have no meaning if less than zero, and some other variables that had skewed distributions, were square-root transformed prior to fitting models.
We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate.
We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone, in combination, and in combination with their interactions. We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] and selected the best model with the combination of the fewest non-significant parameters and lowest AIC~c~.
A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
Mt. Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established (J. Nitta, pers. obs.).

@@

# Do some special formatting for reporting rounded MPDfunc vs MNTDfunc P-val
# - combine into a single value if same, or report each different value with 
# "respectively" if different
if(
  magrittr::equals(
    t_test_print %>%
      filter(var == "mpd.obs.z.func") %>%
      pull(p),
    t_test_print %>%
      filter(var == "mntd.obs.z.func") %>%
      pull(p)
  )
) {
  p_mpd_func_mntd_func <- get_mod_result(t_test_print, "mpd.obs.z.func", "p")
  both <- "both "
} else {
  p_mpd_func_mntd_func <- paste0(
    get_mod_result(t_test_print, "mpd.obs.z.func", "p"),
    " and ",
    get_mod_result(t_test_print, "mntd.obs.z.func", "p"),
    ", respectively"
  )
  both <- ""
}
```

@@

Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
Spatial autocorrelation was detected in the linear model for SLA (`r s_table("linear-fit")`), so a spatial generalized mixed model was used instead (see Methods). 
Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities.
Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).

>>>>

# Load analysis results
loadd(list = c(
  "best_fit_div_models",
  "climate_ancova_results",
  "climate_plot",
  "climate_raw",
  "climate_select",
  "combined_comm_div_plots",
  "combined_cwm_plots",
  "comm",
  "correlated_evo_test",
  "correlated_evo_test_gemmae",
  "daily_mean_climate",
  "div_el_model_summaries",
  "div_t_test_results",
  "div_metrics_all",
  "fern_traits",
  "importance_heatmap",
  "moorea_climate",
  "pca_plot",
  "pca_results",
  "phy",
  "phylosig_binary_traits",
  "phylosig_cont_traits",
  "pic_results",
  "species_list",
  "traits_with_tree",
  "spatial_sla_model"),
  cache = canopy_cache)

@@

We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, 'climatic variables').
Variables relating to VPD, which have no meaning if less than zero, and some other variables that had skewed distributions, were square-root transformed prior to fitting models.
We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate.
We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone, in combination, and in combination with their interactions. We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] for each model and retained models within 2 AIC~c~ of the lowest (\ie, 2 delta AIC~c~). From these selected we the model with the combination of the fewest non-significant parameters and lowest AIC~c~ (for VPD, we the selected the model producing the smallest amount of negative predicted values).
A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
Mt. Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established (J. Nitta, pers. obs.).

@@

# Do some special formatting for reporting rounded MPDfunc vs MNTDfunc P-val
# - combine into a single value if same, or report each different value with 
# "respectively" if different
if(
  magrittr::equals(
    t_test_print %>%
      filter(var == "mpd.obs.z.func") %>%
      pull(p),
    t_test_print %>%
      filter(var == "mntd.obs.z.func") %>%
      pull(p)
  )
) {
  p_mpd_func_mntd_func <- get_mod_result(t_test_print, "mpd.obs.z.func", "p")
  both <- "both "
} else {
  p_mpd_func_mntd_func <- paste0(
    get_mod_result(t_test_print, "mpd.obs.z.func", "p"),
    " and ",
    get_mod_result(t_test_print, "mntd.obs.z.func", "p"),
    ", respectively"
  )
  both <- ""
}

# Extract model summary values for SLA
sla_model <-
spatial_sla_model %>%
  pull(model) %>%
  magrittr::extract2(1)

sla_parameters <- spaMM::get_ranPars(sla_model)

sla_nu <- sla_parameters[["corrPars"]][["1"]][["nu"]] %>% round(2)
sla_rho <- sla_parameters[["corrPars"]][["1"]][["rho"]] %>% round(2)
sla_AIC <- AIC(sla_model)[[1]] %>% round(2)
```

@@

Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
Spatial autocorrelation was detected in the linear model for SLA (`r s_table("linear-fit")`), so a spatial generalized mixed model was used instead (see Methods). 
Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (spatial generalized linear mixed model, nu = `r sla_nu`, rho = `r sla_rho`, AIC = `r sla_AIC`; `r s_table("spatial-sla")`).
Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).
